High-accuracy neutron diffusion calculations based on integral transport theory

### 2019-12-10 docs/EPJP2019/Gross-EPJP-2019.tex
revision of the theory part by Tomatis

\begin{figure}[htbp] 
	\centering
	\begin{tikzpicture}
	\tikzstyle{every node}=[font=\footnotesize]
	\draw[thick] (0,0) -- (3,0);
	\node[anchor=south] at (3.5cm,-5pt) {$\ldots$};
	\draw[thick] (4,0) -- (7,0);
	\node[anchor=south] at (7.5cm,-5pt) {$\ldots$};
	\draw[thick,->] (8,0) -- (11,0) node[anchor=west] {$x$};

@@

Two one-dimensional test cases are considered, homogeneous and heterogeneous. The reference solutions are calculated using a Python3 discrete ordinates $S_N$ code with $N=16$~\cite{Lewis-1984}. The diffusion solutions and the Ronen iterations are produced using original codes still developed in Python3. All tolerances on relative residuals are set to 1E-6 in the iterative solving schemes and no (over/under)-relaxation is done. The initial diffusion coefficient is set to $D=1/(3\sigma_{tr}) = 1/(3\sigma)$ since we only consider here isotropic scattering. Anisotropic scattering cases are deferred to future studies. Void boundary conditions are used at the edges of the slab, which are implemented for the diffusion solver in the form of a group-dependent extrapolated distance, i.e., $d_g = 2.13 D_g$, see~\sxn{sec:BC}.

@@

Table~\ref{tab:dx_drho} shows the reactivity differences as a function of mesh refinement using standard diffusion (with extrapolated length) and the Ronen method. The reactivity difference is given by $(\Delta\rho = 1/k_\texttt{ref} - 1/k_\texttt{D/RM})\times 10^5$, where $k_D$ and $k_{RM}$ are the multiplication factors of standard diffusion and of the Ronen method, respectively. The reference $k_\texttt{ref}$ value from S\tsub{16} is 0.744417 ($I = 400$).

@@

A comparison of the fast and thermal fluxes, as calculated by the reference $S_N$ code, the RM code, and the standard multigroup diffusion without RM correction ($D_0$), is shown in Fig.~\ref{fig:slab-fluxes} for the half slab. The deviations (in [\%]) of the RM-corrected flux and of the standard diffusion flux from the reference $S_N$ flux are also shown in Fig.~\ref{fig:slab-fluxes-dev}. The corresponding Ronen method correction factors $\delta D$ are shown in Fig.~\ref{fig:slab-RM-corr-factor}.

@@

\begin{figure}[htbp]
	% convergence
	\centering
	\includegraphics[width=0.4\linewidth]{convergence.pdf}
	\caption{Convergence of the flux (max deviation) and of the criticality
		eigenvalue.}
	\label{fig:conv}
\end{figure}

The RM correction terms ($\delta D$) along the slab are shown in Fig.~\ref{fig:slab-RM-corr-factor}. It is clear that the diffusion correction assumes zero values at the slab center and exhibit non-trivial behavior near the boundary. The convergence of the flux (max deviation) and of the criticality eigenvalue are shown in Fig.~\ref{fig:conv}. While the eigenvalue converges fast within a few iterations, the flux exhibit slow convergence in general with higher differences against the reference $S_{16}$ solution near the boundaries. The spatial flux convergence between two successive Ronen iterations is shown in Fig.~\ref{fig:conv2} for the left half slab, according to $\Delta \phi = (\phi^{k-1}-\phi^k)/\phi^{k-1}$ [\%]. The spatial flux convergence of the Ronen iterations with respect to the reference solution for the left half slab is shown in Fig.~\ref{fig:conv3} instead.

% convergence
\begin{figure}[htbp]
	\centering
	\includegraphics[width=0.48\linewidth]{flux_deviation_half_it_1.pdf}
	\includegraphics[width=0.48\linewidth]{flux_deviation_half_it_2.pdf}
	\caption{The spatial flux convergence between two successive Ronen iterations for the left half slab, according to $\Delta\phi = (\phi^{(k-1)}-\phi^{(k)})/\phi^{(k-1)}$ [\%].}
	\label{fig:conv2}
\end{figure}

% convergence
\begin{figure}[hbtp]
	\centering
	\includegraphics[width=0.48\linewidth]{flux_deviation_half_it_1_sn.pdf}
	\includegraphics[width=0.48\linewidth]{flux_deviation_half_it_2_sn.pdf}
	\caption{The spatial flux convergence of the Ronen iterations with respect to the reference $S_N$ solution for the left half slab.}
	\label{fig:conv3}
\end{figure}

Notice the change of sign in the flux convergence $\Delta \phi$ in Fig.~\ref{fig:conv2} between the center (negative) and the boundary (positive) of the slab. The initial diffusion solution is smoother than the transport one. Hence, it can be seen in Fig.~\ref{fig:slab-fluxes} that the diffusion solution underestimates the peak flux at the center of the slab and overestimates the flux near the boundary. Ronen iterations ``push'' the diffusion solution towards the transport one by increasing the flux level at the center of the slab and decreasing the flux level at the boundary. Hence, $\phi^{(k-1)}$ is smaller at the center and higher at the boundary compared to $\phi^{(k)}$. This effect caused by Ronen iterations produces less smooth flux shape with steeper gradients comparing to the diffusion solution.
%
% ---------------------------------------------------
\subsection{Heterogeneous case}
\label{subsec:heterog}

>>>>

\begin{figure}[htbp!] 
	\centering
	\begin{tikzpicture}
	\tikzstyle{every node}=[font=\footnotesize]
	\draw[thick] (0,0) -- (3,0);
	\node[anchor=south] at (3.5cm,-5pt) {$\ldots$};
	\draw[thick] (4,0) -- (7,0);
	\node[anchor=south] at (7.5cm,-5pt) {$\ldots$};
	\draw[thick,->] (8,0) -- (11,0) node[anchor=west] {$x$};

@@

Two one-dimensional test cases are considered, homogeneous and heterogeneous. The reference solutions are calculated using a Python3 discrete ordinates $S_N$ code with $N=16$~\cite{Lewis-1984}. The diffusion solutions and the Ronen iterations are produced using original codes developed in Python3. All tolerances on relative residuals are set to 1E-6 in the iterative solving schemes and no (over/under)-relaxation is done. The initial diffusion coefficient is set to $D=(3\sigma_{tr})^{-1}=(3\sigma)^{-1}$ since we only consider here anisotropic scattering. Anisotropic scattering cases are deferred to future studies. Void boundary conditions are used at the edges of the slab, which are implemented for the diffusion solver in the form of a group-dependent extrapolated distance, i.e., $d_g = 2.13 D_g$.

@@

Table~\ref{tab:dx_drho} shows the reactivity differences as a function of mesh refinement using standard diffusion (with extrapolated length) and the Ronen method. The reactivity difference is given by $(\Delta\rho = 1/k_\texttt{ref} - 1/k_\texttt{D/RM})\times 10^5$, where $k_D$ and $k_{RM}$ are the multiplication factors of standard diffusion and the Ronen method, respectively. The reference $k_\texttt{ref}$ value from S\tsub{16} is 0.744417 ($I = 400$).

@@

A comparison of the fast and thermal fluxes, as calculated by the reference $S_N$ code, the RM code, and a standard multigroup diffusion without RM correction ($D_0$), is shown in Fig.~\ref{fig:slab-fluxes} for the half slab. The deviations (in [\%]) of the RM-corrected flux and the standard diffusion ($D_0$) flux from the reference $S_N$ flux is also shown in Fig.~\ref{fig:slab-fluxes-dev}. The corresponding Ronen method correction factors $\delta D$ are shown in Fig.~\ref{fig:slab-RM-corr-factor}.

@@

\begin{figure}[htbp]
	% convergence
	\centering
	\includegraphics[width=0.4\linewidth]{convergence.pdf}
	\caption{Convergence of the flux (max deviation) and the criticality
		eigenvalue.}
	\label{fig:conv}
\end{figure}

The RM correction terms ($\delta D$) along the plane are shown in Fig.~\ref{fig:slab-RM-corr-factor}. It is clear that the diffusion correction assumes zero values at the slab center and exhibit non-trivial behavior near the boundary. The convergence of the flux (max deviation) and the criticality eigenvalue are shown in Fig.~\ref{fig:conv}. While the eigenvalue converges within a few iterations, the flux exhibit slow convergence near the boundaries. The spatial flux convergence between two successive Ronen iterations is shown in Fig.~\ref{fig:conv2} for the left half slab, according to $\Delta\phi = (\phi^{k-1}-\phi^k)/\phi^{k-1}$ [\%]. The spatial flux convergence of the Ronen iterations with respect to the reference $S_N$ solution for the left half slab is shown in Fig.~\ref{fig:conv3}.

% convergence
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.48\linewidth]{flux_deviation_half_it_1.pdf}
	\includegraphics[width=0.48\linewidth]{flux_deviation_half_it_2.pdf}
	\caption{The spatial flux convergence between two successive Ronen iterations for the left half slab, according to $\Delta\phi = (\phi^{(k-1)}-\phi^{(k)})/\phi^{(k-1)}$ [\%].}
	\label{fig:conv2}
\end{figure}

% convergence
\begin{figure}[htbp!]
	\centering
	\includegraphics[width=0.48\linewidth]{flux_deviation_half_it_1_sn.pdf}
	\includegraphics[width=0.48\linewidth]{flux_deviation_half_it_2_sn.pdf}
	\caption{The spatial flux convergence of the Ronen iterations with respect to the reference $S_N$ solution for the left half slab.}
	\label{fig:conv3}
\end{figure}

Notice the change of sign in the flux convergence $\Delta \phi$ in Fig.~\ref{fig:conv2} between the center (negative) and the boundary (positive) of the slab. The initial diffusion solution is smoother than the transport one. Hence, as can be seen in Fig.~\ref{fig:slab-fluxes}, the diffusion solution underestimates the peak flux at the center of the slab and overestimates the flux near the boundary. The Ronen iterations ``push'' the diffusion solution towards the transport one by increasing the flux level at the center of the slab and decreasing the flux level at the boundary. Hence, $\phi^{(k-1)}$ is smaller in the center and larger at the boundary comparing to $\phi^{(k)}$. This effect of the Ronen iterations produces less smooth flux shape with steeper gradients comparing to the diffusion solution.
%
% ---------------------------------------------------
\subsection{Heterogeneous case}
\label{subsec:heterog}
